---
title: "Assignment 1: Oakland's TOD"
author: "Alex Nelms"
date: "9/8/2021"
output: github_document
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(tigris_class = "sf")
```

```{r setup_package, warning = FALSE, message = FALSE}
library(tidyverse)
library(tidycensus)
library(sf)
library(tmap)
library(ggrepel)
library(tigris)

set.seed(717)

mapTheme <- function(base_size = 12) {
  theme(
    text = element_text( color = "black"),
    plot.title = element_text(size = 16,colour = "black"),
    plot.subtitle=element_text(face="italic"),
    plot.caption=element_text(hjust=0),
    axis.ticks = element_blank(),
    panel.background = element_blank(),axis.title = element_blank(),
    axis.text = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(colour = "black", fill=NA, size=2),
    strip.text.x = element_text(size = 14))
}

plotTheme <- function(base_size = 12) {
  theme(
    text = element_text( color = "black"),
    plot.title = element_text(size = 16,colour = "black"),
    plot.subtitle = element_text(face="italic"),
    plot.caption = element_text(hjust=0),
    axis.ticks = element_blank(),
    
    panel.background = element_blank(),
    panel.grid.major = element_line("grey80", size = 0.1),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(colour = "black", fill=NA, size=2),
    strip.background = element_rect(fill = "grey80", color = "white"),
    strip.text = element_text(size=12),
    
    axis.title = element_text(size=12),
    axis.text = element_text(size=10),
    
    
    plot.background = element_blank(),
    
    legend.background = element_blank(),
    legend.title = element_text(colour = "black", face = "italic"),
    legend.text = element_text(colour = "black", face = "italic"),
    
    strip.text.x = element_text(size = 14)
  )
}

```

# Bay Area Rapid Transit (BART) in the East Bay

### A Transit Oriented Development Study

------------------------------------------------------------------------

## 1. Data Wrangling

### 1.A Transit Routes & Stops

Transit Routes & Stops are for the Bay Area Regional Transit, the San Francisco Bay Area's hybrid subway-commuter rail.
Looking at this transportation system allows us to focus on both urban cetners and their immediate suburbs.
The data was sourced from the regional [Metropolitan Transportation Commission](https://opendata.mtc.ca.gov/ "MTC Open Data Portal").

```{r 1A_transit_import, echo=FALSE, include = FALSE}

stop_path = "https://services3.arcgis.com/i2dkYWmb4wHvYPda/arcgis/rest/services/transitstops_existing_planned_2021/FeatureServer/0/query?where=1%3D1&outFields=*&outSR=2227&f=geojson"

route_path = "https://services3.arcgis.com/i2dkYWmb4wHvYPda/arcgis/rest/services/transitroutes_01_2020/FeatureServer/0/query?where=1%3D1&outFields=*&outSR=2227&f=geojson"

BART.stops =
  # from the MTC API, who sourced it from BART
  st_read(stop_path) %>%
  # only look at BART & weekday routes
  filter(agency_id == "BA", route_s_nm!="Blue-Sun", route_s_nm!='Beige') %>%
  mutate(
    route_s_nm = ifelse(
      route_s_nm=="Blue-Wkd/Sat", "Blue", 
      route_s_nm)
  ) %>%
  st_transform(., 2227)
# Projection is (NAD83) California State Plan, Zone 2 (in FEET)
# EPSG:2226
CA_crs = st_crs(BART.stops)
#cat(CA_crs$input, ', EPSG:', CA_crs$epsg)

BART.routes = 
  st_read(route_path) %>%
  filter(agency_id == "BA", route_s_nm!="Blue-Sun") %>%
  mutate(
    route_s_nm = ifelse(
      route_s_nm=="Blue-Wkd/Sat", "Blue", 
      route_s_nm)
  ) %>%
  st_transform(., CA_crs)
```

### 1.A 2 Prepare Transit & East Bay City Boundary Data

Narrowing down Census Tracts & Transit stations to the urban area of the Eastern San Francisco Bay Area *"East Bay"*: Oakland, Berkeley, and the immediate suburban cities of Alameda, Albany, Emeryville, and Albany.*

```{r 1A_transit_prep, include = FALSE}

# PULL CITY BOUNDARIES FROM REGIONAL MTC OPEN DATA
city_path = 'https://services3.arcgis.com/i2dkYWmb4wHvYPda/arcgis/rest/services/region_jurisdiction_clp/FeatureServer/0/query?where=1%3D1&outFields=jurname,fipco&outSR=2227&f=geojson'
EB.cities = st_read(city_path) 

# NARROW AREAS TO MAJOR & MINOR CITIES
major_EB = c('Berkeley', 'Oakland')
minor_EB = c('Alameda', 'Albany', 'Emeryville', 'Piedmont')
is_EB = c(minor_EB, major_EB)
EB.cities = EB.cities[match(is_EB, EB.cities$jurname), ] %>%
  select( -fipco) %>%
  rename(
    city = jurname
  ) %>%
  st_transform(., CA_crs)
EB.cities$size = 'minor'
EB.cities[EB.cities$city %in% major_EB, 'size'] = 'major'

# NARROW STOPS TO .5 MILE OUTSIDE THESE CITIES
BART.stops =
  BART.stops %>%
  # .5 mile buffer + 1 foot
  st_filter(., st_buffer(st_union(EB.cities), dist=5280*.5+1)) %>%
  group_by(stop_nm,stop_id) %>%
  st_drop_geometry(.) %>%
  summarise(
    route_count = n_distinct(route_s_nm), # get routes used
    routes      = list(unique(route_s_nm))
  ) %>%
  merge(., BART.stops[,'stop_id'], by='stop_id') %>%
  distinct() %>% # remove duplicates
  st_as_sf() 

# NARROW STOPS TO THESE CITIES 
BART.routes = st_intersection(BART.routes, st_buffer(st_union(EB.cities),5280*.5))

# CREATE CITY LABELS
EB.labels = 
  EB.cities %>%
  st_centroid(EB.cities) %>%
  rename(
    label = city
  ) %>%
  mutate(
       lon = map_dbl(geometry, ~st_centroid(.x)[[1]]),
       lat = map_dbl(geometry, ~st_centroid(.x)[[2]])
     )

```

### 1.B American Community Survey Census Tracts (2009 & 2019)
*Loading Census Tracts & Variables*

```{r 1B.1_Tracts}

# ACS VARIABLES TO FOCUS ON
acs_var = c(
  "B01001_001", # total pop
  # HOUSING
  "B19013_001", # median household income
  "B25001_001", # housing units
  "B25058_001", # Median Contract rent
  "B25071_001", # median rent as percent of household income
  "B07003_001", # total
  "B07003_004", # total same house 1 year
  # TRAVEL
  "B08135_001", # aggregate time to work
  "B08135_010", # aggregate time to work over 60 mins
  # DEMOGRAPHIC
  "B01001_010", # Male Age    22 to 24
  "B01001_011", # Male Age    25 to 29
  "B01001_012", # Male Age    30 to 34
  "B01001_034", # Female Age  22 to 24
  "B01001_035", # Female Age  25 to 29
  "B01001_036"  # Female Age  30 to 34
  # THIS SECTION OF MY CODE IS DEDICATED TO THE VARIABLES THAT COULD HAVE BEEN STUDIED IF TIDY CENSUS DIDNT HAVE ERRORS WITH THESE TABLES
  # Rest In Pieces
  ## "B08201_001", # household size
  ## "B08201_002", # households with not vehicle
  ## "B24114_001", # total occupations
  ## "B24114_068"  # total software developers
  )

# FUNCTION TO PULL EACH ACS YEAR'S CENSUS TRACTS
get_acs_year <- function(acs_year) {
  # TIDY CENSUS PULL
  acs_tracts = 
      get_acs(
        geography = "tract", 
        variables = acs_var, 
        year=acs_year, state='CA', county='Alameda',
        geometry=T) %>%
      st_transform(CA_crs)
  acs_tracts = 
    acs_tracts %>%
    dplyr::select( -NAME, -moe) %>%
    spread(variable, estimate)
  # GROUP BY GEOID TO FIX MULTIPOLYGON ISSUE
  acs_tracts =
    acs_tracts %>%
    group_by(GEOID) %>%
    summarise(
      year          = acs_year,
      tot_pop       = first(B01001_001), 
      med_HH_inc    = first(B19013_001),
      med_rent      = first(B25058_001),
      tot_units     = first(B25001_001),
      pct_60m_commute = first(B08135_010) / first(B08135_001), 
      tot_22_34yo   = first(B01001_010) + first(B01001_011) + first(B01001_012) 
        + first(B01001_034) + first(B01001_035) + first(B01001_036),
      pct_22_34yo   = tot_22_34yo/first(B01001_001),
      tot_moved     = first(B07003_001) - first(B07003_004),
      pct_moved     = tot_moved/first(B07003_001),
      pct_med_rent  = first(B25071_001),
      geometry      = st_union(geometry) # turn to multipoly
      )
  # REMOVE RAW ACS FIELDS
  non_raw_cols = 
    colnames(acs_tracts)[!startsWith(colnames(acs_tracts), "B")]
  acs_tracts = acs_tracts[non_raw_cols]
  
  return (acs_tracts)
}

# ADD ACS EMPLOYMENT FIELDS THAT TIDYCENSUS WOULDN'T IMPORT
get_acs_subinfo = function(acs_year){   
  # local csv
  directory = 'C:/Users/nelms/OneDrive/Penn/MUSA-508/Files'
  filename = gsub(" ","",paste('census', acs_year, '_employment.csv'))
  
  focus_year = read.csv(file=file.path(directory, filename))
  focus_year = 
    focus_year %>%
    dplyr::select( -NAME, -moe, -X) %>%
    spread(variable, estimate)
  focus_year = 
    focus_year %>%
    mutate(
      # total employed people (16yr +)
      tot_loc_emp = C24050_001, 
      # total employed people in information sector 
      # includes (software, tech, media)
      tot_loc_emp_info = C24050_008, 
      year = acs_year
    )
  # format GEOID like tidycensus
  focus_year$GEOID = 
    lapply(as.character(focus_year$GEOID), 
           function(x) paste('0',x,sep=''))
  # remove raw census columns
  non_raw_cols = 
    colnames(focus_year)[!startsWith(colnames(focus_year), "C")]
  focus_year = focus_year[non_raw_cols]
  
  return(focus_year)
}

tracts09 = get_acs_year(2009)
tracts19 = get_acs_year(2019)

alltracts = st_as_sf(rbind(tracts09, tracts19))

acs_info = rbind(get_acs_subinfo(2009), get_acs_subinfo(2019))

alltracts = st_as_sf(merge(alltracts, acs_info, 
                           by = c('GEOID','year')))

```

### 1.B 2 Narrowing to the East Bay
*Select Tracts & Routes, Clip by Water, Calculate Variables*

```{r 1B.2_EastBay}

# CREATE ERASE FUNCTION
st_erase <- function(x, y) {
  st_difference(x, st_union(y))}

# ERASE WATER AREAS FROM CENSUS TRACTS
EB.water =
  area_water("CA", "Alameda", class = "sf") %>%
  st_transform(., CA_crs)
alltracts <- st_erase(alltracts, EB.water)

# GET CENTROIDS IN EAST BAY
alltracts.centroid =
  alltracts[c('GEOID','geometry')] %>%
  # centroid with largest poly to avoid water in city shapes
  st_centroid(alltracts, of_largest_polygon = TRUE) %>%
  st_join(., EB.cities, left=TRUE)
alltracts.centroid = 
  alltracts.centroid[alltracts.centroid$city %in% is_EB,]
# SELECT TRACTS BY EAST BAY
center_geoms = alltracts.centroid$GEOID
alltracts = alltracts[alltracts$GEOID %in% center_geoms,]

# ERASE CITIES BY WATER
EB.cities <- st_erase(EB.cities, EB.water)

# CALCULATE NEW COLUMNS BY WEIGHTING & LQ
alltracts =
  alltracts %>%
  mutate(
    # Location Quotient of Information Industry Employees
    lq_emp_info = 
      # ratio of info employees to total for tracts "local"
      (tot_loc_emp_info/tot_loc_emp) / 
      # ratio of info o total for East Bay "regional"
      (sum(tot_loc_emp_info) / sum(tot_loc_emp)),
    # area of tract that's not water
    land_area   = st_area(geometry),
    # population density
    dens_ppl    = tot_pop/land_area,
    # housing unit density
    dens_units  = tot_units /land_area,
    # median rent (in 2009) adjusted by 2019 inflation
    med_rent_inf = 
      ifelse(year == "2009", med_rent * 1.19, med_rent),
    # median rent (in 2009) adjusted by 2019 inflation
    med_HH_inc_inf = 
      ifelse(year == "2009", med_HH_inc * 1.19, med_HH_inc)
  )

# PLOT CURRENT DATA
ggplot() +
  # base tracts
  geom_sf(data=alltracts, size=.5,
          color='grey80', fill='grey90') +
  # cities
  geom_sf(data=EB.cities, fill='transparent', 
          color=alpha('grey50',.5),
          lwd=1, linetype = "dashed") + 
  # centroids
  geom_sf(data=alltracts.centroid, color='grey50', size=.5) +
  # BART routes
  geom_sf(data=BART.routes, color=BART.routes$route_s_nm, lwd=1) + 
  # BART stops
  geom_sf(data=BART.stops, fill='white',
          color='grey30', shape=21) + 
  # labels
  geom_text(
    data=EB.labels, check_overlap=TRUE,
    size = 2, fontface='bold', color='grey20',
    aes(x=lon,y=lat, label=label)) + 
  theme(axis.title=element_blank(), 
       axis.text=element_blank(),
       axis.ticks=element_blank()) + 
  mapTheme()

at = st_as_sf(alltracts)

```

## 1.B 3 Finding TOD Areas
*Buffering Transit (BART) Stops then Select Census Tracts by Area in TOD*

```{r 1B.3_TOD}

alltracts = st_as_sf(at)

# .5 MILE BUFFERS OF BART STOPS
BART.buffers <- 
  rbind(
    st_buffer(BART.stops, 5280*.5) %>%
      mutate(Legend = "Buffer") %>%
      dplyr::select(Legend),
    st_union(st_buffer(BART.stops, 5280*.5)) %>%
      st_sf() %>%
      mutate(Legend = "Unioned Buffer"))
buffer = filter(BART.buffers, Legend=="Unioned Buffer")

# CLIP TRACTS BY BUFFER
clip = 
  st_intersection(buffer, alltracts[alltracts$year == 2019,]) %>%
    dplyr::select(GEOID, land_area) %>%
    mutate(
      # find area that is in the buffer
      clip_area = st_area(geometry),
      # percent of buffer area to total
      pct_in_tod = as.vector(clip_area / land_area),
      Selection_Type = "Clip")
clipped = clip[c("GEOID","pct_in_tod","clip_area")]
clipped$geometry = NULL

# GET TOTAL & PERCENT LAND IN CLIP
alltracts = 
  merge(alltracts, clipped, by = "GEOID", all.x = TRUE)
alltracts[is.na(alltracts$clip_area), 'clip_area'] = 0
alltracts[is.na(alltracts$pct_in_tod), 'pct_in_tod'] = 0
alltracts$clip_area = as.vector(alltracts$clip_area)

# GET OTHER SELECTIONS FOR COMPARING
selection = 
  alltracts[buffer,] %>%
    dplyr::select(GEOID, pct_in_tod) %>%
    mutate(
      Selection_Type = "Spatial Selection")
selectCentroids =
  alltracts.centroid[buffer,] %>%
    st_drop_geometry() %>%
    left_join(dplyr::select(alltracts, GEOID, pct_in_tod)) %>%
    st_sf() %>%
    mutate(
      Selection_Type = "Select by Centroids"
      )

# NARROW DOWN TOD TRACTS BASED ON CLIPPED STOPS BUFFER AREA (% and TOT)
tod_tracts = as.data.frame(alltracts[
  # select tracts that had 50% of their land in .5mi stops buffer
  (alltracts$pct_in_tod>=.5)|
  # select tracts that had:
  ( ### at least 20% of their land in the buffer, and
    (alltracts$pct_in_tod<.5)&
    (alltracts$pct_in_tod>.20)&
    ### at least .01 sq.miles (278,784 )
    (alltracts$clip_area>278784)
  ),])[,'GEOID']

alltracts$TOD = FALSE
alltracts[alltracts$GEOID %in% tod_tracts, "TOD"] = TRUE


ggplot() +
  geom_sf(data=clip, aes(fill=pct_in_tod)) +
  geom_sf(data=selection, aes(fill=pct_in_tod)) +
  geom_sf(data=selectCentroids, aes(fill=pct_in_tod)) +
  geom_sf(data=BART.buffers, fill='transparent', color='red') + 
  geom_sf(
    data=alltracts[(alltracts$pct_in_tod>=.5)|((alltracts$pct_in_tod<.5) & (alltracts$pct_in_tod>.20) & (alltracts$clip_area>278784)), ], fill='transparent', color='black'
          ) + 
  geom_sf(
    data=alltracts[alltracts$GEOID=='06001400200',], fill='yellow'
  )+
  scale_fill_steps2(
    low = "red",
    mid = "white",
    midpoint=.5,
    high = "blue",
    name='% within TOD buffer'
  ) + 
  facet_wrap(~Selection_Type) + 
  mapTheme()




```

### 1.D Crime Data

Crime

```{r 1D_Crime}

api_key = "ZfuhpSK1tX9XxWicYM7ew1lHSPYCF9mI2MnKZdBd"

tracts09 %>% 
  group_by(GEOID)
  summarise(st_union(geometry))

t09 <- tracts09 %>%
  group_by(GEOID, year) %>%
  summarise(geometry = st_union(.)) %>%
  summarise_at(vars(-geometry), first)

```

------------------------------------------------------------------------

## 2 Four Time & Space Maps

***Comparing TOD & non-TOD Tracts between 2000 & 2019***

Median Rent, [Second], [Third], [Fourth]

### 2.A Population Density 
*People per Square Foot (of Land Area)*

```{r 2A_variable1}

qN <- function(variable, buckets) {as.factor(ntile(variable, buckets))}

ggplot()+
  geom_sf(data = EB)+
  geom_sf(allTracts, aes(fill = qN(population, 5))) +
  geom_sf(data = BART.buffers, fill = "transparent", color = "red")+
  scale_fill_manual(values = palette5,
                    labels = qBr(allTracts.group, "population"),
                    name = "Population\n(Quintile Breaks, TOD in Red)") +
  labs(title = "Total Population 2010-2018", subtitle = "Boston, MA", caption="Figure 3.1") +
  facet_wrap(~year)+
  mapTheme() + 
  theme(plot.title = element_text(size=15))

```

### 2.B Percent of Median Household Income spent on Rent

```{r 2B_variable2}

```

### 2.C Percent of People who have moved in the Last Year

```{r 2C_variable3}

```

### 2.D Location Quotient of People employed in the Information Industry
*Ratio of the amount of Information industry employees (e.g. software, technology, & media companies)*
*to the total amount of employees — both in the Census Tract & Region (East Bay)*

```{r 2D_variable4}

```

------------------------------------------------------------------------

## 3 Bar Plot

*One grouped bar plot making these same comparisons*

```{r 3_barplot}

```

------------------------------------------------------------------------

## 4 Table

*One table making these same comparisons*

```{r 4_table}

```

------------------------------------------------------------------------

## 5 Graduated Symbol Maps

*Two Graduated Symbol Maps within 0.5 miles transit stations*

### 5.B Map of Population within ½ Mile of BART Stations

```{r 5A_pop}

```

### 5.B Map of Rent within ½ Mile of BART Stations

```{r 5B_rent}

```

------------------------------------------------------------------------

## 6 Geom Line Plots

***Rent & Crime compared to the Distance to BART stations***

```{r 6_geom_line}

```

------------------------------------------------------------------------

## 7 Policy Brief

------------------------------------------------------------------------

## Load data from {tidycensus}

```{r acs_vars, cache = TRUE, message = FALSE, warning = FALSE, results=FALSE}
acs_vars <- c(
    "B01001_001",    # ACS total Pop stimat
    "B15003_001",
    "B25002_001",     # stimat of total housing units
    "B25002_003",     # Numbr of vacant housing units
    "B19013_001",    # Mdian HH Incom ($)
    "B02001_002",  # Popl dscribing thmslvs as "whit alon"
    "B07010_001",     # stimat!!Total Individuals
    "B07010_002",  # stimat!!Total:!!No incom
    "B07010_011", # stimat!!Total:!!With incom:!!$75,000 or mor
    "B07010_023",   # stimat!!Total:!!Movd within sam county:
    "B07010_034",  # stimat!!Total:!!Movd from diffrnt county:
    "B07010_045",   # stimat!!Total:!!Movd from diffrnt country:
    "B07010_012",  # Estimate!!Total:!!Same house 1 year ago:
    "B15003_023",  #Estimate!!Total:!!Disc jockeys, except radio
    "B15003_024",  #Estimate!!Total:!!Disc jockeys, except radio
    "B15003_025"  #Estimate!!Total:!!Disc jockeys, except radio

              ) 

myTracts <- c(
    "42101007400", 
    "42101007300", 
    "42101008000", 
    "42101007800", 
    "42101007700",
    "42101008000",
    "42101007900",
    "42101008802", # just west of UPENN
    #"42101036900", # UPENN, hospitals, woodlands
    #"42101008801", # UPENN
    "42101008702",
    "42101008701",
    "42101008601"
              )


acsTractsPHL.2019.sf <- get_acs(geography = "tract",
                             year = 2019,
                             variables = acs_vars,
                             geometry = TRUE,
                             state  = "PA",
                             county = "Philadelphia",
                             output = "wide") %>%
  dplyr::select (GEOID, NAME, all_of(paste0(acs_vars,"E"))) %>%
  rename (
      total_pop.2019 = B01001_001E,    # ACS total Pop estimate
      total_25yrpop.2019 = B15003_001E,
      total_HU.2019 = B25002_001E,     # Estimate of total housing units
      total_vacant.2019 = B25002_003E,     # Number of vacant housing units
      med_HH_income.2019 = B19013_001E,    # Median HH Income ($)
      total_white.2019 = B02001_002E,  # People describing themselves as "white alone"
      total_indi.2019 = B07010_001E,     # Estimate!!Total Individuals
      total_no_income.2019 = B07010_002E,  # Estimate!!Total:!!No income
      total_income_75k.2019 = B07010_011E, # Estimate!!Total:!!With income:!!$75,000 or more
      total_moved_incnty.2019 = B07010_023E,   # Estimate!!Total:!!Moved within same county:
      total_moved_outcnty.2019 = B07010_034E,  # Estimate!!Total:!!Moved from different county:
      total_moved_outusa.2019 = B07010_045E,   # Estimate!!Total:!!Moved from different country:
      total_same_house_1yr.2019 = B07010_012E,  # Estimate!!Total:!!Same house 1 year ago:
      total_masters.2019 = B15003_023E
  ) %>%
  mutate(
    Neighborhood = ifelse(GEOID %in% myTracts,
                               "CLARK PARK",
                               "REST OF PHILADELPHIA"),
    moved_1yr_PCT.2019 = (1 - total_same_house_1yr.2019 / total_indi.2019 )*100,
    masters_PCT.2019 = ((total_masters.2019 + B15003_024E+ B15003_025E) / total_25yrpop.2019 )*100
    )
```

## Transform to WGS84 with {sf}

```{r}
acsTractsPHL.2019.sf <- acsTractsPHL.2019.sf %>% 
  st_transform(crs = "EPSG:4326")
```

## Plot with {ggplot2}

```{r ggplot_geom_sf, warning = FALSE, echo = FALSE}
max_pct = max(acsTractsPHL.2019.sf$moved_1yr_PCT.2019[is.finite(acsTractsPHL.2019.sf$moved_1yr_PCT.2019)])

ggplot()+
  # POLYGON color 
  geom_sf(data = acsTractsPHL.2019.sf, 
          aes(fill = moved_1yr_PCT.2019),
          color = "transparent")+
  # blank on top
  geom_sf(data = acsTractsPHL.2019.sf %>%
            filter(Neighborhood == "CLARK PARK") %>%
            st_union(),
            color = "black",
            size=1, #linetype="longdash",
            fill = "transparent"
          ) + 
  scale_fill_gradient2(
      low="green4", mid="white", high="red4", midpoint=20,
      breaks = c(0,20,40,60), labels = c("0%","20%","40%","60%"),
      name="% of New Residents")+
  guides(fill = guide_colourbar(barheight = 10))+
  labs(
    title = "Percentage of individuals who have moved in the last year",
    subtitle = "Focus on the Clark Park neighborhood") +
  theme(plot.title = element_text(hjust = 0.5), axis.title.x=element_blank(), 
        axis.text.x=element_blank(), axis.ticks.x=element_blank(),
        axis.title.y=element_blank(), axis.text.y=element_blank(), 
        axis.ticks.y=element_blank(),
        panel.background = element_rect(fill='gray'))
```

\# )

\#brewer.pal(3, name="Dark2")

```{r ggplot_geom_sf, warning = FALSE, echo = FALSE}

library(RColorBrewer)


ggplot()+
  #POLYGON
  geom_sf(data = acsTractsPHL.2019.sf,
          fill = "white",
          color = "grey75"
          )+
  # POINT
  geom_sf(data = acsTractsPHL.2019.sf  %>%
          st_centroid(),                # to centroid
          shape = 21, color='grey10',
          aes(
            size = masters_PCT.2019,    # set point size by column value
            fill = masters_PCT.2019     # set point fill ('color') by column value
          )) +
  # Scale for the Point Size
  scale_size_continuous(
          range = c(0,6),             # range of point sizes
          breaks = c(0,20,40,60),       # value breaks/bins of point size
          labels = c("0 to 19 %", "20 to 39 %", "40 to 59%", "60+ %"), 
                                        # legend labels
          name="% of People over 25yrs with a \nMaster's, Doctoral, or Professional Degree"
          )+
  scale_fill_stepsn(                    # binned scale of point fill
          colors = RColorBrewer::brewer.pal( # pulling ColorBrewer list of colors
                      4,                # amount of bins
                      name="BuPu"       # ColorBrewer palette
                      ),
          breaks = c(0,20,40,60),       # value breaks/bins of point size
          guide = FALSE                 # remove legend (so we can have one combo)
          )+
    labs(
          title = "Percentage of Adults who earned \n a Master's Degree or Higher",
          ) +
    theme(
          axis.title=element_blank(), 
          axis.text=element_blank(), axis.ticks=element_blank(), # remove ticks
          panel.background = element_rect(fill='gray')
          )+
    guides(
          size = guide_legend(
              override.aes = list(fill = brewer.pal(4, name="BuPu"))
          ))                            # add the ColorBrewer list to the legend
```
